---
title: "Kids with Congenital Heart Defects: Results Report"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: false
    latex_engine: xelatex
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
params:
  group: "simple"
  height: 150
  weight: 45
  bmi: 20
  conf_level: 95
  height_range: !r c(150, 180)
  bmi_range: !r c(18.5, 25)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
source("classes.R")
source("methods.R")
source("models.R")
source("plotting.R")

group_label <- switch(params$group,
  simple  = "Simple defects (ASD, VSD, CoA, LVOT obstruction)",
  moderate = "Moderate complex defects (TOF, TGA)",
  fontan  = "Univentricular defects with Fontan circulation"
)
group_names <- c(simple = "Simple", moderate = "Moderate", fontan = "Fontan")
conf_level <- params$conf_level / 100
```

## Patient Parameters

```{r params-table}
knitr::kable(
  data.frame(
    Parameter = c(
      "Diagnostic Group", "Height", "Weight", "BMI", "Confidence Level"
    ),
    Value = c(
      group_label,
      paste(params$height, "cm"),
      paste(params$weight, "kg"),
      sprintf("%.1f kg/m\u00b2", params$bmi),
      paste0(params$conf_level, "%")
    )
  ),
  col.names = c("Parameter", "Value"),
  booktabs = TRUE
)
```

## Results

The values below are predicted cardiopulmonary exercise test (CPET) endpoints
for the given diagnostic group and patient covariates. Confidence intervals are
calculated using a normal approximation.

```{r compute-results}
grp <- get(params$group)

person_male   <- person(sex = 1, height = params$height, bmi = params$bmi)
person_female <- person(sex = 0, height = params$height, bmi = params$bmi)

get_ci <- function(fun) {
  list(
    male   = fun(grp, person_male,   conf_level),
    female = fun(grp, person_female, conf_level)
  )
}

vo2_ml_min_r     <- get_ci(grp$vo2_ml_min)
vo2_ml_kg_min_r  <- get_ci(grp$vo2_ml_kg_min)
heart_rate_r     <- get_ci(grp$heart_rate)
ventilation_r    <- get_ci(grp$ventilation)
oxygen_pulse_r   <- get_ci(grp$oxygen_pulse)
ve_vco2_slope_r  <- get_ci(grp$ve_vco2_slope)
breathing_freq_r <- get_ci(grp$breathing_frequency)

format_ci <- function(mat, digits = 2) {
  if (is.null(mat) || any(is.na(mat[1, 2:3]))) return(NA_character_)
  paste(
    formatC(mat[1, 2], format = "f", digits = digits),
    "-",
    formatC(mat[1, 3], format = "f", digits = digits)
  )
}

metrics <- c(
  "Absolute VO\u2082 (ml/min)",
  "VO\u2082 relative to body mass (ml/kg/min)",
  "Heart rate (BPM)",
  "Ventilation (l/min)",
  "Oxygen pulse (ml/beat)",
  "Ventilatory efficiency (VE/VCO\u2082 slope)",
  "Breathing frequency (breaths/min)"
)

value_male <- c(
  sprintf("%.1f", vo2_ml_min_r$male[1, 1]),
  sprintf("%.2f", vo2_ml_kg_min_r$male[1, 1]),
  sprintf("%.1f", heart_rate_r$male[1, 1]),
  sprintf("%.2f", ventilation_r$male[1, 1]),
  sprintf("%.2f", oxygen_pulse_r$male[1, 1]),
  sprintf("%.2f", ve_vco2_slope_r$male[1, 1]),
  sprintf("%.2f", breathing_freq_r$male[1, 1])
)
value_female <- c(
  sprintf("%.1f", vo2_ml_min_r$female[1, 1]),
  sprintf("%.2f", vo2_ml_kg_min_r$female[1, 1]),
  sprintf("%.1f", heart_rate_r$female[1, 1]),
  sprintf("%.2f", ventilation_r$female[1, 1]),
  sprintf("%.2f", oxygen_pulse_r$female[1, 1]),
  sprintf("%.2f", ve_vco2_slope_r$female[1, 1]),
  sprintf("%.2f", breathing_freq_r$female[1, 1])
)
ci_male <- c(
  format_ci(vo2_ml_min_r$male,    1),
  format_ci(vo2_ml_kg_min_r$male, 2),
  format_ci(heart_rate_r$male,    1),
  format_ci(ventilation_r$male,   2),
  format_ci(oxygen_pulse_r$male,  2),
  NA_character_,
  format_ci(breathing_freq_r$male, 2)
)
ci_female <- c(
  format_ci(vo2_ml_min_r$female,    1),
  format_ci(vo2_ml_kg_min_r$female, 2),
  format_ci(heart_rate_r$female,    1),
  format_ci(ventilation_r$female,   2),
  format_ci(oxygen_pulse_r$female,  2),
  NA_character_,
  format_ci(breathing_freq_r$female, 2)
)
```

```{r results-table}
ci_col <- sprintf("%d%% CI", params$conf_level)
df <- data.frame(
  Metric      = metrics,
  "Value Boy" = value_male,
  CI_Boy      = ci_male,
  "Value Girl" = value_female,
  CI_Girl     = ci_female,
  check.names = FALSE
)
names(df)[3] <- paste0(ci_col, " Boy")
names(df)[5] <- paste0(ci_col, " Girl")
knitr::kable(df, booktabs = TRUE, caption = "Predicted CPET endpoints")
```

---

*Generated by Kids with Congenital Heart Defects app, Oslo University Hospital,
Department of Paediatric Cardiology.*

*Reference: Klungerbo V, Hirth A, Fredriksen PM, Holst R, Edvardsen E,
Holmstrom H, Moller T. Reference models for individualized assessment of
cardiorespiratory fitness in children and adolescents with congenital heart
disease: a retrospective multicentre study. Eur J Pediatr. 2025 Jun
26;184(7):450. doi: 10.1007/s00431-025-06270-x.*
